---
title: "Databases and Queries in MySQL"
format: pdf
editor: visual
author: Ankur Malik, Haotian Xia, Haoming Deng, Jennifer Park, Russel Liu 
---

## ABSTRACT:

summary:

## DATA:

Write about data set...

# 

## WHAT ARE Queries, SQL, MySQL:

**What is a Query? What is a SQL? What is MySQL?**

A **query** is a statement which enables us to extract data or information from a particular database table. We are writing a statement or command which allows us to extract a specific piece of information from our database. Now you may wonder what a database is? A **database** is the collection of structured data stored within a computer system. We use database management systems (*DBMS*) to control a database.

**SQL** is an acronym for *standard query language*. SQL is considered to be a programming language which can be used to manipulate, query, access, and control a database or table.

In this project, we are focusing on a specific kind of SQL called MySQL**. MySQL** is defined a relational database management system based on SQL. It differs from other versions of SQL primarily based on syntax, functionality, and purpose.

## MySQL Functions:

**Function:** WHERE

**Purpose:** It is used to filter records based on a specific condition

```{r}
# Selecting an instant from the day database wehre the temp is 0.3441670 and limit it to 2 outputs 
dbGetQuery(mydb, "SELECT instant FROM day where temp = 0.3441670 LIMIT 2" ) 
```

**Function:** ORDER BY

**Purpose:** It is used to sort the result-set in ascending or descending order.

```{r}
# Selecting the temperature from the day database and ordering our ouput by season in Ascending order with a limit of 3 ouputs 
dbGetQuery(mydb, "SELECT temp FROM day ORDER BY season ASC LIMIT 3" ) # Ascending Order
# Selecting the temperature from the day database and ordering our ouput by season in Descending order with a limit of 3 ouputs 
dbGetQuery(mydb, "SELECT temp FROM day ORDER BY season DESC LIMIT 3" ) # Descending Order
```

**Function:** LIMIT

**Purpose:** It is used to select a limited number of records.

```{r}
#Select the date from the day database where the tempeature is greater than 0 and limiting it to 5 outputs
dbGetQuery(mydb, "SELECT dteday FROM day where temp > 0 LIMIT 5" )  
```

**Function:** LEFT JOIN

**Purpose:** Left Join returns all records from the left table (table1), and the matching records from the right table (table2). The result is 0 records from the right side, if there is no match.

```{r}
# Selecting windspeed from the day database and casual from the hour database and left joining all the attributes on season from the day and hour database and limiting it to 2 outputs
dbGetQuery(mydb, 'SELECT day.windspeed, hour.casual FROM day LEFT JOIN hour ON day.season = hour.season LIMIT 2')
```

**Function:** RIGHT JOIN

**Purpose:** Right Join returns all records from the right table (table2), and the matching records from the left table (table1). The result is 0 records from the left side, if there is no match.

```{r}
# Selecting windspeed from the day database and casual from the hour database and right joining all the attributes on season from the day and hour database and limiting it to 2 outputs
dbGetQuery(mydb, 'SELECT day.windspeed, hour.casual FROM day RIGHT JOIN hour ON day.season = hour.season LIMIT 2')
```

**Function:** INNER JOIN

**Purpose:** Inner Join selects records that have matching values in both tables.

```{r}
# Selecting windspeed from the day database and casual from the hour database and inner joining all the attributes on season from the day and hour database and limiting it to 2 outputs
dbGetQuery(mydb, 'SELECT day.windspeed, hour.casual FROM day INNER JOIN hour ON day.season = hour.season LIMIT 2')
```

**Function:** OUTER JOIN/ FULL JOIN

**Purpose:** Outer Join keyword returns all records when there is a match in left (table1) or right (table2) table records.

```{r}
# dbGetQuery(mydb, 'SELECT day.season, hour.casual FROM day FULL JOIN hour ON day.season = hour.season LIMIT 5')
# Selecting all the attributes from day and hour database where there is a match
dbGetQuery(mydb, "SELECT day.windspeed, hour.casual FROM day LEFT JOIN hour ON day.season = hour.season UNION 
           SELECT day.windspeed, hour.casual FROM day RIGHT JOIN hour ON day.season = hour.season LIMIT 5")
```

**Function:** UNION

**Purpose:** We use the union operator to combine the result-set of two or more SELECT statements.

```{r}
# We are selecting cnt from the day database and also selecting temperature from the hour database and using union to combine both select statements with a limit of 2 outputs
dbGetQuery(mydb, "SELECT day.cnt FROM day UNION SELECT temp FROM hour LIMIT 2")
```

**Function:** IN

**Purpose:** The in operator allows you to specify multiple values in a where clause or shorthand for multiple or conditions

```{r}
# Selecting attributes from the day database where season in the select season from the day database and limit output to 5     
dbGetQuery(mydb, "SELECT* FROM day WHERE season IN (SELECT season FROM day) LIMIT 5")
```

**Function:** LIKE

**Purpose:** The like operator is used in a where clause to search for a specified pattern in a column. We use the percent operator (%) to represent 0, 1, or multiple characters. The underscore ( \_ ) represents a single character.

```{r}
# Selecting temp and season from the day database where the date... and limit to 4 outputs
dbGetQuery(mydb, "SELECT temp,season FROM day WHERE dteday LIKE '%2' LIMIT 4") # End in 2
dbGetQuery(mydb, "SELECT temp,season FROM day WHERE dteday LIKE '2%' LIMIT 4") # Start with 2 
dbGetQuery(mydb, "SELECT temp,season FROM day WHERE dteday LIKE '%2%' LIMIT 4") # Has 2 in it 
dbGetQuery(mydb, "SELECT temp,season FROM day WHERE dteday LIKE '2%2' LIMIT 4") # Start with 2 and ends in 2

```

**Function:** BETWEEN

**Purpose:** The between operator selects values within a given range. The values can be numbers, text, or dates.

```{r}
#Select the date and atemp from day where the temp is between 0 and 1 and limit to 5 outputs
dbGetQuery(mydb, "SELECT dteday, atemp FROM day where temp BETWEEN 0 AND 1 LIMIT 5" )  

```

**Function:** GROUP BY

**Purpose:** The group by statement groups rows that have the same values into summary rows

```{r}
# Select the number of season from day database and group by date the output and order by the count of the season in descending order and limit to 2 outputs
dbGetQuery(mydb, "SELECT COUNT(season) FROM day GROUP BY dteday ORDER BY COUNT(season) DESC LIMIT 2")
```

**Function:** HAVING

**Purpose:** The having clause was added to SQL because the where keyword cannot be used with aggregate functions.

```{r}
# Select the date from the day database and group by date having the count of the temperatures being greater than 0.2 and limit to 5 outputs
dbGetQuery(mydb, "SELECT dteday FROM day GROUP BY dteday HAVING COUNT(temp) > 0.2 LIMIT 5" ) 
```

**Function:** CASE

**Purpose:** expression goes through conditions and returns a value when the first condition is met (like an if-then-else statement). So, once a condition is true, it will stop reading and return the result. If no conditions are true, it returns the value in the else clause.

```{r}
# Select windspeed and when the windspeed is less than 0.1 output low else Null
dbGetQuery(mydb, "SELECT windspeed, CASE 
            WHEN windspeed < 0.1 THEN 'LOW'
            ELSE NULL
            END AS CHECK_LOWSPEED FROM day")
```

**Function:** DISTINCT

**Purpose:** The distinct statement is used to return only distinct (different) values

```{r}
# Select distinct date from day where the temp is 0.3441670
dbGetQuery(mydb, "SELECT DISTINCT(dteday) FROM day where temp = 0.3441670" ) 
dbGetQuery(mydb, "SELECT DISTINCT(windspeed) FROM day where temp = 0.1" ) # Value Does not Exist
```

## Works Cited:

1.  Colburn, Rafe. Special Edition Using SQL. Que, 2000.
2.  Röhm, Uwe, et al. "SQL for Data Scientists: Designing SQL Tutorials for Scalable Online Teaching." Proceedings of the VLDB Endowment, vol. 13, no. 12, 2020, pp. 2989--92, https://doi.org/10.14778/3415478.3415526.
3.  Guo, Aibo, et al. "ER-SQL: Learning Enhanced Representation for Text-to-SQL Using Table Contents." Neurocomputing (Amsterdam), vol. 465, 2021, pp. 359--70, https://doi.org/10.1016/j.neucom.2021.08.134.
4.  Guo, Aibo, et al. "ER-SQL: Learning Enhanced Representation for Text-to-SQL Using Table Contents." Neurocomputing (Amsterdam), vol. 465, 2021, pp. 359--70, https://doi.org/10.1016/j.neucom.2021.08.134.
5.  Reese, George., et al. Managing and Using MySQL. 2nd ed., O'Reilly, 2002.

## 

\
