---
title: "Haotian'part_w_scope/target_people"
output: pdf_document
date: "2022-12-08"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## We are intruducing a tutorial for people who want to play with database in R 
Goal: Make people feel confortable in basic programing with database and extract the data they want
Target people (group): people who do not have database programing expeirence 
scope: SQL and dbplyr in R

SQL is one of the most polupar database langauge for people to extract data information. However, it has it own syntax and sometimes hard to memorize and learn. In order to adapt people have different coding background, we are also introducing dbplyr as a subsititution way to help you extract data from the dataset. dbplyr is a package in R, and its syntax like most r functions. The purpose for dbplyr is to automatically generate SQL for you so that you’re not forced to use it. 

However, dbplyr doesn’t do everything and is not as powerful as SQL. Therefore, this projct will spend most of time on SQL and show you how dbplyr works in some cases. 

After reading both dbplyr part and sql part, you will know how easy to grab data from dataset and will form a basis for you to use them successfully in real projects. 

Let's begin with the SQL part.


```{r sql part}
setwd('~/Desktop/PSTAT197/vignette-database-configuration/')
library(RMariaDB)
library(DBI)
df_day <- read.csv("data/day.csv", header=TRUE)
head(df_day)

df_hour <-read.csv("data/hour.csv", header=TRUE)
head(df_hour)



# list the tables available in bikes
con <- dbConnect(RMariaDB::MariaDB(), 
                 user = 'root', 
                 password = 'shiatsing2',
                 host = 'localhost')
mydb <- dbConnect(RMariaDB::MariaDB(), 
                  user = 'root', 
                  password = 'shiatsing2',
                  host = 'localhost',
                  dbname = 'bikes')
dbListTables(mydb)


dbReadTable(mydb, 'hour')
dbReadTable(mydb, 'day')

# Count ALL: How many row we have
dbGetQuery(mydb, "SELECT COUNT(*) FROM day" )

# Count Distinct weekday 
dbGetQuery(mydb, "SELECT COUNT(DISTINCT (weekday)) FROM day" )

# we want to see the average windspeed among all data

dbGetQuery(mydb, "SELECT AVG(windspeed) AS AVERAGE_SPEED FROM day" )

# we want to see the sum of casual users
dbGetQuery(mydb, "SELECT SUM(casual) AS TOTAL_NUM_CASUAL_USER FROM day")

# we want to see the max speed of wind
dbGetQuery(mydb, "SELECT MAX(windspeed) AS MAX_WINDSPEED FROM day")

# we want to see the min speed of wind
dbGetQuery(mydb, "SELECT MIN(windspeed) AS MIN_WINDSPEED FROM day")

# We want to concat days and hours as one column

dbGetQuery(mydb, "SELECT CONCAT(dteday,' Hour: ',hr) FROM hour")

# We want to see how many characters composes the date
dbGetQuery(mydb, "SELECT DISTINCT(LENGTH(dteday)) AS length_of_char_date FROM hour")

#WE want to firstly create windspeed breakdown to two part low and high. Then we want to 
# show how to use lower function to make all spring be lower case
dbGetQuery(mydb, "SELECT LOWER(SPEED_BREAKDOWN) FROM (SELECT windspeed, CASE 
            WHEN windspeed < 0.1 THEN 'LOW'
            ELSE 'HIGH'
            END AS SPEED_BREAKDOWN FROM day) AS tmp")

#  We want to show how to use upper function to make all spring be upper case
dbGetQuery(mydb, "SELECT UPPER(SPEED_BREAKDOWN) FROM (SELECT windspeed, CASE 
            WHEN windspeed < 0.1 THEN 'low'
            ELSE 'high'
            END AS SPEED_BREAKDOWN FROM day) AS tmp")
```

Now, Let's talk about dbplyr.

```{r dbplyr}
library(dbplyr)
library(dplyr)
#CONNECT TO DATABASE
con <- dbConnect(RMariaDB::MariaDB(), 
                  user = 'root', 
                  password = 'shiatsing2',
                  host = 'localhost',
                  dbname = 'bikes')

#CONNECT TO TABLE
data_sample<- con %>% tbl("day")


# SELECT differenct columns 
# SELECT 4 cloumns together. In Sql we have to write down the four columns' name seperately. But in dbplyr, we can use : to select a range of columns.
data_sample %>% select(season:holiday)

# We can use filter to subsititude where condition in sql. Here we are showing data that has wind speed greater than 0.4
data_sample %>% filter(windspeed > 0.4)

# We can also do the same SQL group by and aggregation task by uing dbplyr group_by function

data_sample %>% group_by(weekday) %>%
  summarise(
    count_day = count(instant)
  )

```

However if we want to realize join function, we need to use show_query() function to identify the join key. In this case, we still need to use SQL syntax in order to do the comlicated correct join. More can be viewed here: https://dbplyr.tidyverse.org/reference/join.tbl_sql.html

So the dbplyr can be used in samll simple aggregation tasks without sql knowledge. However, once multiple dataset includes, we still need to know SQL in order to finish more complicated tasks. 
