---
title: "Russell_Section2_Numeric"
author: "Russell Liu"
date: "2022-12-08"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

Set up environment:

```{r}
library(RMariaDB)
library(DBI)

df_day <- read.csv("data/day.csv", header=TRUE)
head(df_day)

df_hour <-read.csv("data/hour.csv", header=TRUE)
head(df_hour)
```

```{r}
# list the tables available in bikes
dbListTables(mydb)
```

Numeric Functions

#rand()

```{r}
dbGetQuery(mydb, "SELECT RAND()" )
```

We can see that a random number from 0 to 1 is chosen.

#round()

The ROUND() function rounds a number to a specified number of decimal
places. For example, we want to round up temperature and windspeed to 3
decimals

```{r}
dbGetQuery(mydb, "SELECT temp,round(temp,3),windspeed, round(windspeed,
3) FROM day" )
```

We can see that temperature and windspeed are round from 7decimals to 3
decimals.

#FLOOR()

The FLOOR() function returns the largest integer value that is smaller
than or equal to a number. We want to see the largest integer value
which is either equal to or less than the given average or minimum
windspeed.

```{r}
dbGetQuery(mydb,"SELECT AVG(windspeed),MIN(windspeed),FLOOR(AVG(windspeed)) AS
Lower_Average FROM day" ) 
dbGetQuery(mydb, "SELECT
AVG(windspeed),MIN(windspeed),FLOOR(MIN(windspeed)) AS Lower_Average
FROM day" )
```

Given the average value of windspeed and minimum value of windspeed, we
see largest integer value that is either equal to or less than the
average and minimum windspeed.

#CEIL()

The CEIL() function returns the smallest integer value that is bigger
than or equal to a number. It is equal to the CEILING() function. we
want to see the smallest integer value which is either greater than or
equal to the given average or minimum number.

```{r}
dbGetQuery(mydb, "SELECT
AVG(windspeed),MIN(windspeed),CEIL(AVG(windspeed)) AS Lower_Average FROM
day" ) 
dbGetQuery(mydb, "SELECT
AVG(windspeed),MIN(windspeed),CEIL(MIN(windspeed)) AS Lower_Average FROM
day" )
```

#Truncate(n,d) with positive d

we want to see atemp keeps 2 decimals and hum only keeps 1 decimals

```{r}
dbGetQuery(mydb, "SELECT atemp, hum,
TRUNCATE(atemp,2), TRUNCATE(hum,1) FROM day" )
```

#Truncate(n,d) with negative d we want to see atem and hum are truncated
two or one digit left to the decimal points.

```{r}
dbGetQuery(mydb,
"SELECT atemp, hum, TRUNCATE(atemp,-2), TRUNCATE(hum,-1) FROM day" )
```

```{r}
dbGetQuery(mydb, "SELECT * FROM hour" )
```
